---
title: "Data Analysis"
author: 'Jonathan Yuquimpo'
date: '01-11-21'
output: html_document
---

```{r include=FALSE}
# Package names
packages <- c('tidyverse', 'readr')
# Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}
# Packages loading
invisible(lapply(packages, library, character.only = TRUE))
rm(packages, installed_packages)

data <- read_csv(
  "data/merged_file.csv",
  col_types = cols(
    cntBalance = col_factor(levels = c("r",
                                       "g")),
    correct = col_logical(),
    rt = col_double(),
    rwdType = col_factor(levels = c("0",
                                    "1", "2")),
    tgtCol = col_skip(),
    tgtDir = col_skip(),
    #tgtPos = col_factor(levels = c("0",
    #                               "1", "2", "3", "4", "5")),
    #colPos = col_factor(levels = c("0",
    #                               "1", "2", "3")),
    ...13 = col_skip()
  ),
  na = "empty"
)

`%!in%` <- Negate(`%in%`)

data <- tibble::as_tibble(data)
data <- filter(data,
               trialType == 'RSVPtest',
               subID %!in%  c('brk','0tt','a0v','c90','rp9','tx3','y95','yz3'),
               rt > 250 #quantile(data$rt, 0.025, na.rm = TRUE)
               ) %>%
  select(subID, correct, rt, rwdType, tgtPos, colPos, colMatch)
data$subID <- as.factor(data$subID)
```

## Raw Data
Snippet of data file used for analysis.  
rt: response time  
rwdType: associated value of colored item (if any)  

0. No value (black) 
1. Low value
2. High value

tgtPos: serial position of target item

0. 1st position
1. 2nd position
2. 3rd position
3. New item (for false alarms)

colPos: serial position of the target item

0. 1st position  
1. 2nd position  
2. 3rd position  
3. New item (for false alarms)  

colMatch: Does the color item match the target item?  

0. color item present  
1. color item present but is not mached with target
2. color item present and is matched with target

Add segment about trial proportions
XXX trials. XX of them were new, 
```{r Data output, echo=FALSE}
print(data, n=20)
```

```{r fig.width=10, include=FALSE}

rwdName = c('Low', 'High', 'None')
rwdArr = c(1, 2, 0)


for (i in 1:2) {
  assign(
    paste0('congruent', rwdName[i]),
    bind_cols(
      filter(data,
             rwdType == rwdArr[i],
             tgtPos == rwdArr[i],
             correct == T) %>%
        group_by(subID, .drop = F) %>%
        summarize(hits = n()) %>%
        select(subID, hits),
      filter(data,
             rwdType == rwdArr[i],
             tgtPos == rwdArr[i],
             correct == F) %>%
        group_by(subID, .drop = F) %>%
        summarize(miss = n(),
                  rt = median(rt)) %>%
        select(miss),
      filter(data,
             rwdType == rwdArr[i],
             tgtPos == 3,
             correct == F) %>%
        group_by(subID, .drop = F) %>%
        summarize(fAlm = n(), #fAlm = false alarm
                  rt = median(rt)) %>%
        select(fAlm),
      filter(data,
             rwdType == rwdArr[i],
             tgtPos == 3,
             correct == T) %>%
        group_by(subID, .drop = F) %>%
        summarize(rejt = n(), #rejt = rejections
                  rt = median(rt)) %>%
        select(rejt)
    ) %>%
      mutate (
        hitSum = hits + miss,
        hitRate = hits / hitSum,
        HR = ifelse(
          hitRate == 1,
          (hits - 0.5) / hitSum,
          ifelse(hitRate == 0,
                 (hits + 0.5) / hitSum,
                 hitRate)
        ),
        fAlmSum = fAlm + rejt,
        fAlmRate = fAlm / fAlmSum,
        FA = ifelse(
          fAlmRate == 1,
          (fAlm - 0.5) / fAlmSum,
          ifelse(fAlmRate == 0,
                 (fAlm + 0.5) / fAlmSum,
                 fAlmRate)
        ),
        dprime = qnorm(HR) - qnorm(FA),
        criterion = -(qnorm(HR) + qnorm(FA)) / 2,
        rwd = as.factor(replicate(n = n(), rwdArr[i]))
      )
  )
}

assign(
  paste0('congruent', rwdName[3]),
  bind_cols(
    filter(data,
           rwdType == rwdArr[3],
           tgtPos ==rwdArr[i],
           correct == T) %>%
      group_by(subID, .drop = F) %>%
      summarize(hits = n()) %>%
      select(subID, hits),
    filter(data,
           rwdType == rwdArr[3],
           tgtPos ==rwdArr[i],
           correct == F) %>%
      group_by(subID, .drop = F) %>%
      summarize(miss = n(),
                rt = median(rt)) %>%
      select(miss),
    filter(data,
           rwdType == rwdArr[3],
           tgtPos == 3,
           correct == F) %>%
      group_by(subID, .drop = F) %>%
      summarize(fAlm = n(), #fAlm = false alarm
                rt = median(rt)) %>%
      select(fAlm),
    filter(data,
           rwdType == rwdArr[3],
           tgtPos == 3,
           correct == T) %>%
      group_by(subID, .drop = F) %>%
      summarize(rejt = n(), #rejt = rejections
                rt = median(rt)) %>%
      select(rejt)
  ) %>%
    mutate (
      hitSum = hits + miss,
      hitRate = hits / hitSum,
      HR = ifelse(
        hitRate == 1,
        (hits - 0.5) / hitSum,
        ifelse(hitRate == 0,
               (hits + 0.5) / hitSum,
               hitRate)
      ),
      fAlmSum = fAlm + rejt,
      fAlmRate = fAlm / fAlmSum,
      FA = ifelse(
        fAlmRate == 1,
        (fAlm - 0.5) / fAlmSum,
        ifelse(fAlmRate == 0,
               (fAlm + 0.5) / fAlmSum,
               fAlmRate)
      ),
      dprime = qnorm(HR) - qnorm(FA),
      criterion = -(qnorm(HR) + qnorm(FA)) / 2,
      rwd = as.factor(replicate(n = n(), rwdArr[3]))
    )
)
```

## Filtered Data
For each participant, the number of hits, misses, false alarms, and rejections is totaled and used to compute hit rate and false alarm rate. The 1/2N correction is then applied to hit and false alarm rate which yields the HR and FA columns. Those values are then used to compute dprime and criterion for each reward condition (rwd=0, 1, or 2)
```{r echo=FALSE, R.options = list(width = 200)}
congruentLow
congruentHigh
congruentNone
```

## Computations for graphs
For each variable, standard error is computed within-subjects using the Rmisc package. [See here for more info](http://www.cookbook-r.com/Graphs/Plotting_means_and_error_bars_(ggplot2)/#understanding-within-subjects-error-bars)

```{r include=FALSE}
library(Rmisc)
#install.packages('Rmisc')
#Hit Rate====
longFormData <- bind_rows(congruentHigh, congruentLow, congruentNone)
hit_rate <-
  summarySEwithin(
    longFormData,
    measurevar = 'HR',
    withinvars = 'rwd',
    idvar = 'subID'
  ) %>%
  mutate(condition = as.factor(c(replicate(3, "Hit Rate"))),
         mean = HR)

false_alarm_rate <-
  summarySEwithin(
    longFormData,
    measurevar = 'FA',
    withinvars = 'rwd',
    idvar = 'subID'
  ) %>%
  mutate(condition = as.factor(c(replicate(3, "False Alarm Rate"))),
         mean = FA)

FA_HR_df <- bind_rows(hit_rate, false_alarm_rate) %>%
  select(rwd,mean,sd,se,ci,condition) %>%
  print()

dprime <-
  summarySEwithin(
    longFormData,
    measurevar = 'dprime',
    withinvars = 'rwd',
    idvar = 'subID'
  )

criterion <-
  summarySEwithin(
    longFormData,
    measurevar = 'criterion',
    withinvars = 'rwd',
    idvar = 'subID'
  )
```

```{r echo=FALSE}
FA_HR_df
dprime
criterion
```

```{r include=FALSE}
width = .8
dodgeWidth = .85
errWidth = .4
errdodgeWidth = .85


HR_FA_bar <- ggplot(FA_HR_df, aes(
  y = mean,
  x = condition,
  fill = factor(rwd),
)) +
  geom_bar(
    stat = "identity",
    width = width,
    position = position_dodge(width = dodgeWidth),
    color = 'black'
  ) +
  
  geom_errorbar(
    aes(ymin = mean - se,
        ymax = mean + se),
    width = errWidth,
    position = position_dodge(width = errdodgeWidth),
    colour = "azure4",
    size = 1
  ) +
  
  coord_cartesian(ylim = c(.1, .9)) +
  
  labs(y = 'Proportion',
       x = 'Reward') +
  
  scale_fill_manual(
    values = c('2' = "red",
               '1' = "green",
               '0' = "black"),
    name = 'Reward Condition',
    labels = c('High', 'Low', 'None')
  ) +
  theme_bw()+
  theme(
    legend.position = 'none',
    axis.text = element_text(size = 24),
    axis.title = element_text(size = 32),
    axis.line = element_line(colour = "black"),
    panel.grid.major.x = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank()
  ) 

dprime_bar <- ggplot(dprime,
                     aes(y = dprime,
                         x = rwd,
                         fill = rwd)) +
  
  geom_bar(
    stat = "identity",
    width = width,
    position = position_dodge(dodgeWidth),
    color = 'black'
  ) +
  
  geom_errorbar(
    aes(ymin = dprime - se,
        ymax = dprime + se),
    width = errWidth,
    position = position_dodge(errdodgeWidth),
    colour = "azure4",
    size = 1
  ) +
  
  #coord_cartesian(ylim = c(1.8, 2.2)) +
  #coord_cartesian(ylim = c(1.25, 2.25)) +
  
  
  labs(y = expression(paste("Discriminability ",italic("d\'"))),
       x = '') +
  
  scale_fill_manual(values = c('2' = "red",
                               '1' = "green",
                               '0' = "black")) +
  
  scale_x_discrete(labels = c(
    "2" = "High",
    "1" = "Low",
    "0" = "Control"
  )) +
  
  
  theme_bw()+
  theme(
    legend.position = 'none',
    axis.text = element_text(size = 24),
    axis.title = element_text(size = 32),
    axis.line = element_line(colour = "black"),
    panel.grid.major.x = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank()
  ) 

criterion_bar <- ggplot(criterion,
                        aes(
                          y = criterion,
                          x = interaction(rwd),
                          group = rwd,
                          fill = rwd
                        )) +
  
  geom_bar(
    stat = "identity",
    width = width,
    position = position_dodge(dodgeWidth),
    color = 'black'
  ) +
  #coord_cartesian(ylim = c(-.25, .2)) +
  
  geom_errorbar(
    aes(ymin = criterion - se,
        ymax = criterion + se),
    width = errWidth,
    position = position_dodge(errdodgeWidth),
    colour = "azure4",
    size = 1
  ) +
  
  labs(y = expression(paste("Criterion ",italic("c"))),
       x = '') +
  
  scale_fill_manual(values = c('2' = "red",
                               '1' = "green",
                               '0' = "black")) +
  
  scale_x_discrete(labels = c(
    "2" = "High",
    "1" = "Low",
    "0" = "Control"
  )) +
  
  
  theme_bw()+
  theme(
    legend.position = 'none',
    axis.text = element_text(size = 24),
    axis.title = element_text(size = 32),
    axis.line = element_line(colour = "black"),
    panel.grid.major.x = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank()
  ) 

```

```{r echo=FALSE}
HR_FA_bar
dprime_bar
criterion_bar
```





























































